---
title: "AsthmaDemo_preprocessing.Rmd"
author: "Tara Eicher"
date: "`r Sys.Date()`"
output: html_document
---

# Set up.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
result_dir <- "~\\asthma_vignette_results"
dir.create(result_dir)
```

# Install the packages.
```{r eval = FALSE}
if(!require("devtools")){
  install.packages("devtools")
}
library("devtools")
if(!require("MultiOmicsGraphPrediction")){
  install_github("ncats/MultiOmicsGraphPrediction")
}
library(MultiOmicsGraphPrediction)
```

# Filter metabolites and genes.

This data includes GACRS gene expression and metabolite abundance levels for 320 patients. The outcome of interest is lung function. We received the following analyte pairs of interest from Rachel Kelly:
- IDO1 and IDO2 and the tryptophan kynurenine pathway. The tryptophan kynurenine pathway includes the following pathways in RaMP:
  - NAD biosynthesis II (from tryptophan)
  - NAD+ biosynthesis
  - Tryptophan catabolism
  - Tryptophan catabolism leading to NAD+ production
  - Tryptophan metabolism
- ORMDL3 and sphingolipids and ceramides.
  - Sphingolipids: 

```{r}
# Genes and metabolites of interest (metabolites are from RaMP).
genes.tryptophan.kynurenine <- c("IDO1", "IDO2")
nad.biosynthesis.ii.from.tryptophan <- c("hmdb:HMDB0000045", "hmdb:HMDB0000148", 
                                         "hmdb:HMDB0000161", "hmdb:HMDB0000232", 
                                         "hmdb:HMDB0000538", "hmdb:HMDB0000641", 
                                         "hmdb:HMDB0000684", "hmdb:HMDB0000929")
nad.plus.biosynthesis <- c("hmdb:HMDB0000232", "hmdb:HMDB0000684", 
                           "hmdb:HMDB0000929", "hmdb:HMDB0001406")
tryptophan.catabolism <- c("hmdb:HMDB0000243", "hmdb:HMDB0000684", 
                           "hmdb:HMDB0000715", "hmdb:HMDB0002302")
tryptophan.catabolism.leading.to.nad.plus.production <- c("hmdb:HMDB0000161", 
                                                          "hmdb:HMDB0000197", 
                                                          "hmdb:HMDB0000232", 
                                                          "hmdb:HMDB0000259", 
                                                          "hmdb:HMDB0000684", 
                                                          "hmdb:HMDB0000715", 
                                                          "hmdb:HMDB0000881", 
                                                          "hmdb:HMDB0000929", 
                                                          "hmdb:HMDB0001123", 
                                                          "hmdb:HMDB0001406")
tryptophan.metabolism <- c("hmdb:HMDB0000148", "hmdb:HMDB0000161", "hmdb:HMDB0000197", 
                           "hmdb:HMDB0000232", "hmdb:HMDB0000259", "hmdb:HMDB0000538", 
                           "hmdb:HMDB0000684", "hmdb:HMDB0000715", "hmdb:HMDB0000881", 
                           "hmdb:HMDB0000929", "hmdb:HMDB0001123")

genes.sphingolipid.ceramide <- c("ORMDL3")
sphingolipids <- c("hmdb:HMDB0000277", "hmdb:HMDB0001348", "hmdb:HMDB0004949",
                   "hmdb:HMDB0004952", "hmdb:HMDB0004953", "hmdb:HMDB0004956",
                   "hmdb:HMDB0010169", "hmdb:HMDB0011697", "hmdb:HMDB0012102",
                   "hmdb:HMDB0012107", "hmdb:HMDB0012097", "hmdb:HMDB0012101",
                   "hmdb:HMDB0012103", "hmdb:HMDB0012104", "hmdb:HMDB0000252",
                   "hmdb:HMDB0000269", "hmdb:HMDB0000277", "hmdb:HMDB0001348",
                   "hmdb:HMDB0004949", "hmdb:HMDB0004952", "hmdb:HMDB0004953",
                   "hmdb:HMDB0004956", "hmdb:HMDB0010169", "hmdb:HMDB0011697",
                   "hmdb:HMDB0012102", "hmdb:HMDB0012107", "hmdb:HMDB0012097",
                   "hmdb:HMDB0012101", "hmdb:HMDB0012103", "hmdb:HMDB0012104")

ceramides <- c("hmdb:HMDB0004949", "hmdb:HMDB0004952", "hmdb:HMDB0004953",
               "hmdb:HMDB0004956", "hmdb:HMDB0004949", "hmdb:HMDB0004952",
               "hmdb:HMDB0004953", "hmdb:HMDB0004956")

# Match lists to actual metabolites present in the database.
metab.fdata <- read.csv("/Volumes/eichertd$/asthma_data/cra/metabfdata.csv")
genes.of.interest <- c(genes.tryptophan.kynurenine, genes.sphingolipid.ceramide)
getMetabIds <- function(idList){
  return(unlist(lapply(idList, function(id){
    retVal <- NA
    whichId <- which(metab.fdata$HMDB_ID == id)
    if(length(whichId) > 0){
      retVal <- metab.fdata$id[whichId]
    }
    return(retVal)
  })))
}
getMetabNames <- function(idList){
  return(make.names(unlist(lapply(idList, function(id){
    retVal <- NA
    whichId <- which(metab.fdata$HMDB_ID == id)
    if(length(whichId) > 0){
      retVal <- metab.fdata$Metabolite[whichId]
    }
    return(retVal)
  }))))
}
nad.biosynthesis.ii.from.tryptophan.ids <- getMetabIds(nad.biosynthesis.ii.from.tryptophan)
nad.plus.biosynthesis.ids <- getMetabIds(nad.plus.biosynthesis)
tryptophan.catabolism.ids <- getMetabIds(tryptophan.catabolism)
tryptophan.catabolism.leading.to.nad.plus.production.ids <- getMetabIds(tryptophan.catabolism.leading.to.nad.plus.production)
tryptophan.metabolism.ids <- getMetabIds(tryptophan.metabolism)
sphingolipids.ids <- getMetabIds(sphingolipids)
ceramides.ids <- getMetabIds(ceramides)

nad.biosynthesis.ii.from.tryptophan.names <- getMetabIds(nad.biosynthesis.ii.from.tryptophan)
print(nad.biosynthesis.ii.from.tryptophan.names)
nad.plus.biosynthesis.names <- getMetabIds(nad.plus.biosynthesis)
print(nad.plus.biosynthesis.names)
tryptophan.catabolism.names <- getMetabIds(tryptophan.catabolism)
print(tryptophan.catabolism.names)
tryptophan.catabolism.leading.to.nad.plus.production.names <- getMetabIds(tryptophan.catabolism.leading.to.nad.plus.production)
print(tryptophan.catabolism.leading.to.nad.plus.production.names)
tryptophan.metabolism.names <- getMetabIds(tryptophan.metabolism)
print(tryptophan.metabolism.names)
sphingolipids.names <- getMetabIds(sphingolipids)
print(sphingolipids.names)
ceramides.names <- getMetabIds(ceramides)
print(ceramides.names)

# Read analyte files.
metabs <- read.csv("/Volumes/eichertd$/asthma_data/cra/named_metabdata.csv", row.names = 1)
genes <- read.csv("/Volumes/eichertd$/asthma_data/cra/transcriptdata.csv", row.names = 1)
pdata <- read.csv("/Volumes/eichertd$/asthma_data/cra/pdata.csv", row.names = 1)
fdata.gene <- read.csv("/Volumes/eichertd$/asthma_data/cra/genefdata.csv", row.names = 1)
row.names(pdata) <- make.names(row.names(pdata))
```

```{r}
# Filter metabolites.
all.ids <- c(nad.biosynthesis.ii.from.tryptophan.ids, nad.plus.biosynthesis.ids,
             tryptophan.catabolism.ids, tryptophan.catabolism.leading.to.nad.plus.production.ids,
             tryptophan.metabolism.ids, sphingolipids.ids, ceramides.ids)
all.ids <- unique(all.ids[which(!is.na(all.ids))])
all.names <- c(nad.biosynthesis.ii.from.tryptophan.names, nad.plus.biosynthesis.names,
             tryptophan.catabolism.names, tryptophan.catabolism.leading.to.nad.plus.production.names,
             tryptophan.metabolism.names, sphingolipids.names, ceramides.names)
all.names <- unique(all.names[which(!is.na(all.names))])
metabs <- metabs[all.ids,]
rownames(metabs) <- all.names
colnames(metabs) <- sample(colnames(metabs), ncol(metabs))
csvtrain <- metabs[,1:254]
csvtest <- metabs[,255:318]
write.csv(csvtrain, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/metabDataAsthma.train.csv")
write.csv(csvtest, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/metabDataAsthma.test.csv")

# Filter genes.
csvtrain <- genes[genes.of.interest,1:254]
csvtest <- genes[genes.of.interest,255:318]
write.csv(csvtrain, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/geneDataAsthma.train.csv")
write.csv(csvtest, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/geneDataAsthma.test.csv")

# Filter pdata.
csvtrain <- pdata[1:254,]
csvtest <- pdata[255:318,]
write.csv(csvtrain, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/pDataAsthma.train.csv")
write.csv(csvtest, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/pDataAsthma.test.csv")

# Filter fdata.
fdataFilt <- metab.fdata[all.ids,]
rownames(fdataFilt) <- make.names(rownames(fdataFilt))
fdataFilt$id <- all.names
write.csv(fdataFilt, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/fDataMetabAsthma.csv")
fdataGenes <- fdata.gene[genes.of.interest,]
write.csv(fdataGenes, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/fDataGeneAsthma.csv")

# Write reference file.
refdata.train <- data.frame(type = c("analyteType2", "analyteType1", "analyteType2MetaData", "analyteType1MetaData", "sampleMetaData"),
                      filenames = c("metabDataAsthma.train.csv", "geneDataAsthma.train.csv", "fDataMetabAsthma.csv", "fDataGeneAsthma.csv", "pDataAsthma.train.csv"))
refdata.test <- data.frame(type = c("analyteType2", "analyteType1", "analyteType2MetaData", "analyteType1MetaData", "sampleMetaData"),
                      filenames = c("metabDataAsthma.test.csv", "geneDataAsthma.test.csv", "fDataMetabAsthma.csv", "fDataGeneAsthma.csv", "pDataAsthma.test.csv"))
write.csv(refdata.train, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/Asthma.train.csv", row.names = FALSE)
write.csv(refdata.test, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/Asthma.test.csv", row.names = FALSE)
```

## Plot Analyte Pair Distribution
```{r}
# List analytes of interest.
metabolites_of_interest <- c(rep("2'-deoxyuridine 5'-triphosphate", 7), rep("citric acid", 11),
                             rep("malic acid", 6), rep("succinate", 9))
genes_of_interest <- c(c("FURIN", "UMPS", "UPP1", "UPP2", "UCK1", "RRM2", "TYMS"),
                       c("CNDP1", "APRT", "SLCO2B1", "CLYBL", "BHMT", "PDE5A", "IL4I1", "HS3ST3A1", "HS3ST3B1", "OR10H3"),
                       c("MDH2", "FH", "ME2", "ME1", "ME3", "MDH1"),
                       c("TMLHE", "ALKBH2", "PHYH", "P4HA2", "P4HA1", "BBOX1", "PLOD1", "PLOD2", "PLOD3", "P3H1"))

# Plot analyte pairs.
for(i in 1:length(metabolites_of_interest)){
  numeric_metab <- as.numeric(metabs[metabolites_of_interest[i],])
  numeric_gene <- as.numeric(genes[genes_of_interest[i],])
  which_neither_na <- intersect(which(!is.na(numeric_metab)), which(!is.na(numeric_gene)))
  str(numeric_gene)
  plot(y = numeric_metab[which_neither_na], 
       x = numeric_gene[which_neither_na],
       xlab = genes_of_interest[i], ylab = metabolites_of_interest[i])
}
```

## Plot Analyte Distribution With Respect to 5-FU
```{r}
metabolites_of_interest <- c("2'-deoxyuridine 5'-triphosphate", "citric acid",
                             "malic acid", "succinate", "cysteine")
genes_of_interest <- c("FURIN", "UMPS", "UPP1", "UPP2", "UCK1", "RRM2", "TYMS",
                       "CNDP1", "APRT", "SLCO2B1", "CLYBL", "BHMT", "PDE5A", "IL4I1", "HS3ST3A1", "HS3ST3B1", "OR10H3",
                       "MDH2", "FH", "ME2", "ME1", "ME3", "MDH1",
                       "TMLHE", "ALKBH2", "PHYH", "P4HA2", "P4HA1", "BBOX1", "PLOD1", "PLOD2", "PLOD3", "P3H1",
                       "E2F1", "TP53", "TNFSF10")

# Plot analyte pairs.
for(i in 1:length(metabolites_of_interest)){
  plot(y = metabs[metabolites_of_interest[i], shared_samps], x = pdata$drug5FU, 
       ylab = metabolites_of_interest[i], xlab = "Mean Tegafur Score")
}
for(i in 1:length(genes_of_interest)){
  plot(y = genes[genes_of_interest[i], shared_samps], x = pdata$drug5FU, 
       ylab = genes_of_interest[i], xlab = "Mean Tegafur Score")
}
```
