---
title: "NCI605FUDemo_preprocessing.Rmd"
author: "Tara Eicher"
date: "`r Sys.Date()`"
output: html_document
---

# Set up.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
result_dir <- "~\\Ensemble_5fu_vignette_results"
dir.create(result_dir)
```

# Install the packages.
```{r eval = FALSE}
if(!require("devtools")){
  install.packages("devtools")
}
library("devtools")
if(!require("MultiOmicsGraphPrediction")){
  install_github("ncats/MultiOmicsGraphPrediction")
}
library(MultiOmicsGraphPrediction)
```

# Filter metabolites and genes.

This data includes NCI-60 gene expression and metabolite abundance levels for 57 cell lines. The outcome of interest is response to 5-FU drugs, as reviewed in the following papers: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3837334/ and https://pubmed.ncbi.nlm.nih.gov/34151400/. From the first paper, we obtained the following list of analytes of interest that are contained within the NCI-60 analyte data:
- dUTP (metabolite)
- FURIN (gene)
- UMPS (gene)
- UPP1 and UPP2 (genes)
- UCK1 (gene)
- RRM2 (genes)
- TYMS (gene)

From the second paper, we obtained the following additional list of analytes. These analytes were shown to be independently altered with respect to 5-FU administration in humans:
- Citric acid (metabolite hmdb:HMDB0000094)
- Malic acid (metabolite hmdb:HMDB0000156)
- Succinate (metabolite hmdb:HMDB0000254)
- Cysteine (metabolite hmdb:HMDB0003417)
- E2F1 (gene ensembl:ENSG00000101412)
- TP53 (gene ensembl:ENSG00000141510)
- TNFSF10 (gene ensembl:ENSG00000121858)

For the second list, we also queried the RaMP GUI (RaMP 2.0, https://rampdb.nih.gov/) for analytes involved in reactions with this list and obtained the following additional analytes:
- For citric acid: CNDP1, APRT, SLCO2B1, CLYBL, BHMT, PDE5A, IL4I1, HS3ST3A1, HS3ST3B1, and OR10H3
- For malic acid: MDH2, FH, ME2, ME1, ME3, and MDH1
- For succinate: TMLHE, ALKBH2, PHYH, P4HA2, P4HA1, BBOX1, PLOD1, PLOD2, PLOD3, and P3H1

5-FU drug response for the NCI-60 cell lines was downloaded from the following link: https://discover.nci.nih.gov/cellminer/downloadquery.do.

## Plot 5-FU Response Distribution
```{r}
# Read and process 5-FU response.
#dir <- system.file("extdata", package="MultiOmicsGraphPrediction", mustWork=TRUE)
dir <- "/Volumes/eichertd$/Ensemble_prediction/inst/extdata"
drugs_5fu <- c("Tegafur", "Capecitabine")
drug_responses_init <- read.csv(paste(dir, "DTP_NCI60_RAW.csv", sep = "/"), header = T, skip = 8)
drug_names <- drug_responses_init$Drug.name
drug_responses_trimmed <- t(drug_responses_init[,10:ncol(drug_responses_init)])
drug_responses <- matrix(as.numeric(drug_responses_trimmed), ncol = ncol(drug_responses_trimmed))
rownames(drug_responses) <- rownames(drug_responses_trimmed)

# Read analyte files.
metabs <- read.csv("/Volumes/eichertd$/IntLIM_vignette_code/IntLIMVignettes/NCI60_data/metabData.csv", row.names = 1)
genes <- read.csv("/Volumes/eichertd$/IntLIM_vignette_code/IntLIMVignettes/NCI60_data/geneData.csv", row.names = 1)
pdata <- read.csv("/Volumes/eichertd$/IntLIM_vignette_code/IntLIMVignettes/NCI60_data/pData.csv", row.names = 1)
fdata <- read.csv("/Volumes/eichertd$/IntLIM_vignette_code/IntLIMVignettes/NCI60_data/fData.metab.csv", row.names = 1)
row.names(pdata) <- make.names(row.names(pdata))

# Take subset of 5-FU data.
rownames(drug_responses) <- sub("^.*?\\.", "", rownames(drug_responses))
drug_responses <- drug_responses[sort(intersect(colnames(metabs), rownames(drug_responses))),]

# Plot a histogram of the drug info.
# We remove column 13 because all but one value is NA.
which_drugs <- which(drug_names %in% drugs_5fu)
drugs_heatmap <- drug_responses[,which_drugs[-13]]
colnames(drugs_heatmap) <- paste(drug_names[which_drugs[-13]], drug_responses_init$Experiment.name[which_drugs[-13]], sep = " - ")
gplots::heatmap.2(drugs_heatmap, na.color = "gray", tracecol = NA, scale = "none", margins = c(10,5), cexCol = 0.75)

# Take the mean of all but the leftmost four data points.
experiments_of_interest <- c("9402AG15", "8905CS23", "9402SA37", "8904SA83", "8904NS27", "9403SA38", "9403AG16", "8904NS20", "9403SS94", "8904SA76")
drugs_heatmap_filtered <- drugs_heatmap[,paste("Tegafur", experiments_of_interest, sep = " - ")]
drug_levels_mean <- rowMeans(drugs_heatmap_filtered, na.rm = TRUE)
drug_levels_median <- unlist(lapply(1:nrow(drugs_heatmap_filtered), function(i){return(median(drugs_heatmap_filtered[i,], na.rm = TRUE))}))
hist(drugs_heatmap_filtered, breaks = 25, xlim = c(0,6))
hist(drug_levels_mean, breaks = 25, xlim = c(0,6))
hist(drug_levels_median, breaks = 25, xlim = c(0,6))

# Since taking the mean does not provide a wide enough range, we take the experiment that has the fewest NA values. In two experiments,
# there are only 4 missing values each.
count_na <- unlist(lapply(1:ncol(drugs_heatmap_filtered), function(i){return(length(which(is.na(drugs_heatmap_filtered[,i]))))}))
least_na_drugs <- drugs_heatmap_filtered[,which(count_na == 4)]
hist(least_na_drugs[,1], breaks = 25, xlim = c(0,6))
hist(least_na_drugs[,2], breaks = 25, xlim = c(0,6))

# The second of these has the best range. We check to see how much the high values correspond to colon cancer cell lines.
colon_cell_lines <- c("COLO.205", "HCC.2998", "HCT.116", "HCT.15", "HT29", "KM12", "SW.620")
hist(least_na_drugs[colon_cell_lines,2], breaks = 25, xlim = c(0,6))

# Finalize the drug measurements we are using and filter the cell lines.
final_drug <- colnames(drugs_heatmap_filtered)[which(count_na == 4)[2]]
final_cell_lines <- rownames(drugs_heatmap_filtered)[which(!is.na(drugs_heatmap_filtered[,final_drug]))]
shared_samps <- sort(intersect(colnames(metabs), colnames(genes)))
shared_samps <- intersect(shared_samps, final_cell_lines)
metabs <- metabs[,shared_samps]
genes <- genes[,shared_samps]
pdata <- pdata[shared_samps,]
pdata$drug5FU <- least_na_drugs[shared_samps, 2]
```

## Plot Analyte Pair Distribution
```{r}
# List analytes of interest.
metabolites_of_interest <- c(rep("2'-deoxyuridine 5'-triphosphate", 7), rep("citric acid", 11),
                             rep("malic acid", 6), rep("succinate", 9))
genes_of_interest <- c(c("FURIN", "UMPS", "UPP1", "UPP2", "UCK1", "RRM2", "TYMS"),
                       c("CNDP1", "APRT", "SLCO2B1", "CLYBL", "BHMT", "PDE5A", "IL4I1", "HS3ST3A1", "HS3ST3B1", "OR10H3"),
                       c("MDH2", "FH", "ME2", "ME1", "ME3", "MDH1"),
                       c("TMLHE", "ALKBH2", "PHYH", "P4HA2", "P4HA1", "BBOX1", "PLOD1", "PLOD2", "PLOD3", "P3H1"))

# Plot analyte pairs.
for(i in 1:length(metabolites_of_interest)){
  numeric_metab <- as.numeric(metabs[metabolites_of_interest[i],])
  numeric_gene <- as.numeric(genes[genes_of_interest[i],])
  which_neither_na <- intersect(which(!is.na(numeric_metab)), which(!is.na(numeric_gene)))
  str(numeric_gene)
  plot(y = numeric_metab[which_neither_na], 
       x = numeric_gene[which_neither_na],
       xlab = genes_of_interest[i], ylab = metabolites_of_interest[i])
}
```

## Plot Analyte Distribution With Respect to 5-FU
```{r}
metabolites_of_interest <- c("2'-deoxyuridine 5'-triphosphate", "citric acid",
                             "malic acid", "succinate", "cysteine")
genes_of_interest <- c("FURIN", "UMPS", "UPP1", "UPP2", "UCK1", "RRM2", "TYMS",
                       "CNDP1", "APRT", "SLCO2B1", "CLYBL", "BHMT", "PDE5A", "IL4I1", "HS3ST3A1", "HS3ST3B1", "OR10H3",
                       "MDH2", "FH", "ME2", "ME1", "ME3", "MDH1",
                       "TMLHE", "ALKBH2", "PHYH", "P4HA2", "P4HA1", "BBOX1", "PLOD1", "PLOD2", "PLOD3", "P3H1",
                       "E2F1", "TP53", "TNFSF10")

# Plot analyte pairs.
for(i in 1:length(metabolites_of_interest)){
  plot(y = metabs[metabolites_of_interest[i], shared_samps], x = pdata$drug5FU, 
       ylab = metabolites_of_interest[i], xlab = "Mean Tegafur Score")
}
for(i in 1:length(genes_of_interest)){
  plot(y = genes[genes_of_interest[i], shared_samps], x = pdata$drug5FU, 
       ylab = genes_of_interest[i], xlab = "Mean Tegafur Score")
}
```

```{r}
# Filter metabolites.
colnames(metabs) <- sample(colnames(metabs), ncol(metabs))
csvtrain <- metabs[metabolites_of_interest,1:40]
csvtest <- metabs[metabolites_of_interest,41:51]
write.csv(csvtrain, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/metabData5FU_NCI60.train.csv")
write.csv(csvtest, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/metabData5FU_NCI60.test.csv")

# Filter genes.
colnames(genes) <- colnames(metabs)
csvtrain <- genes[genes_of_interest,1:40]
csvtest <- genes[genes_of_interest,41:51]
write.csv(csvtrain, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/geneData5FU_NCI60.train.csv")
write.csv(csvtest, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/geneData5FU_NCI60.test.csv")

# Filter pdata.
rownames(pdata) <- colnames(metabs)
csvtrain <- pdata[1:40,]
csvtest <- pdata[41:51,]
write.csv(csvtrain, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/pData5FU_NCI60.train.csv")
write.csv(csvtest, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/pData5FU_NCI60.test.csv")

# Filter fdata.
fdataFilt <- fdata[metabolites_of_interest,]
rownames(fdataFilt) <- make.names(rownames(fdataFilt))
write.csv(fdataFilt, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/fDataMetab5FU_NCI60.csv")
fdataGenes <- data.frame(id = genes_of_interest, databaseId = paste("gene_symbol", genes_of_interest, sep = ":"))
rownames(fdataGenes) <- make.names(genes_of_interest)
write.csv(fdataGenes, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/fDataGene5FU_NCI60.csv")

# Write reference file.
refdata.train <- data.frame(type = c("analyteType2", "analyteType1", "analyteType2MetaData", "analyteType1MetaData", "sampleMetaData"),
                      filenames = c("metabData5FU_NCI60.train.csv", "geneData5FU_NCI60.train.csv", "fDataMetab5FU_NCI60.csv", "fDataGene5FU_NCI60.csv", "pData5FU_NCI60.train.csv"))
refdata.test <- data.frame(type = c("analyteType2", "analyteType1", "analyteType2MetaData", "analyteType1MetaData", "sampleMetaData"),
                      filenames = c("metabData5FU_NCI60.test.csv", "geneData5FU_NCI60.test.csv", "fDataMetab5FU_NCI60.csv", "fDataGene5FU_NCI60.csv", "pData5FU_NCI60.test.csv"))
write.csv(refdata.train, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/nci60_5FU.train.csv", row.names = FALSE)
write.csv(refdata.test, "/Volumes/eichertd$/Ensemble_prediction/inst/extdata/nci60_5FU.test.csv", row.names = FALSE)
```
