
---
title: "NCI-60 5-FU Demo"
author: "Tara Eicher"
date: "4/01/2022"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{MultiOmicsGraphPrediction NCI60 5FU}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# Set up.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
result_dir <- "~\\Ensemble_5fu_vignette_results"
dir.create(result_dir)
```

# Install the packages.
```{r eval = FALSE}
if(!require("devtools")){
  install.packages("devtools")
}
library("devtools")
if(!require("MultiOmicsGraphPrediction")){
  install_github("ncats/MultiOmicsGraphPrediction")
}
library(MultiOmicsGraphPrediction)
if(!require("IntLIM")){
  install_github("ncats/IntLIM")
}
library(IntLIM)
```

# Import data files.
Next, we read in the new data.
```{r}
#dir <- system.file("extdata", package="MultiOmicsGraphPrediction", mustWork=TRUE)
dir <- "/Volumes/eichertd$/Ensemble_prediction/inst/extdata"
# Import training data.
csvfileTrain <- file.path(dir, "nci60_5FU.train.csv")
inputDataTrain <- IntLIM::ReadData(csvfileTrain, class.feat = list(drugscore = "numeric",
                                                                   cancertype = "factor",
                                                                   drug5FU = "numeric"))

# Import testing data.
csvfileTest <- file.path(dir, "nci60_5FU.test.csv")
inputDataTest <- IntLIM::ReadData(csvfileTest, class.feat = list(drugscore = "numeric",
                                                                   cancertype = "factor",
                                                                   drug5FU = "numeric"))
```

# Run IntLIM
```{r}
myres <- IntLIM::RunIntLim(inputData = inputDataTrain,stype="drug5FU",
                                 save.covar.pvals = TRUE, 
                                 outcome = 2, suppressWarnings = TRUE,
                                 independent.var.type = 1, 
                                 continuous = TRUE, covar = "cancertype")
# IntLIM::DistPvalues(myres, adjusted = FALSE)
# IntLIM::DistRSquared(myres)
# IntLIM::InteractionCoefficientGraph(inputResults = myres, 
#                                     interactionCoeffPercentile = 0.9,
#                                     percentageToPlot = 1)
# IntLIM::pvalCoefVolcano(inputResults = myres, inputData = inputDataTrain,
#                         pvalcutoff = 0.99)
# plot(density(myres@interaction.pvalues), xlab = "Interaction Term Unadjusted p-value", main = "")
# plot(density(myres@interaction.adj.pvalues), xlab = "Interaction Term Adjusted p-value", main = "")
# plot(density(myres@interaction.coefficients), xlab = "Interaction Term Coefficient", main = "")
# plot(density(myres@model.rsquared), xlab = "Model R-squared", main = "")
```

# Filter Results and Create Co-Regulation Graph
Test the number of pairs returned for different levels of filtering.
```{r}
# Compute all percentiles of R^2 cutoffs.
myres.all <- IntLIM::ProcessResults(inputResults = myres, inputData = inputDataTrain, 
                                       pvalcutoff = 1, rsquaredCutoff = 0, interactionCoeffPercentile = 0)
#R2cutoffs <- unlist(lapply(seq(0, 90, by = 10), function(R2percentile){
#  return(quantile(myres.all$rsquared, R2percentile / 100))
#}))
R2cutoffs <- seq(0, 90, by = 10) / 100
# Create a list of all possible R^2 and interaction coefficient combinations.
ziplist <- paste(unlist(lapply(seq(0,90,by=10), function(num){return(rep(num, 10))})), 
                 rep(R2cutoffs, 10))

# For all combinations, find the pairs.
pairs <- lapply(ziplist, function(cutoff){
  cutoffVals <- as.numeric(strsplit(cutoff, " ")[[1]])
  return(IntLIM::ProcessResults(inputResults = myres, inputData = inputDataTrain,
                                       pvalcutoff = 1, rsquaredCutoff = cutoffVals[2], 
                                interactionCoeffPercentile = cutoffVals[1] / 100))
})
names(pairs) <- ziplist

# For all combinations, return the pair counts.
counts <- unlist(lapply(pairs, function(p){
  return(nrow(p))
}))
countMatrix <- matrix(counts, ncol = 10)
colnames(countMatrix) <- seq(0,90,by=10)
rownames(countMatrix) <- R2cutoffs

# Save file.
#write.csv(countMatrix, file=paste(getwd(), "countMatrix.csv", sep = "/"))
```

# Filter Results and Create Co-Regulation Graph
Filter by R^2 only.
```{r}
myres.r2.density <- density(myres.all$rsquared)
peak.r2 <- myres.r2.density$x[which.max(myres.r2.density$y)]
myres.r2 <- IntLIM::ProcessResults(inputResults = myres, inputData = inputDataTrain, 
                                       pvalcutoff = 1, rsquaredCutoff = peak.r2)
# Build graph.
coreg.r2 <- MultiOmicsGraphPrediction::BuildCoRegulationGraph(myres.r2)
#MultiOmicsGraphPrediction::PlotCoRegulationGraph(coreg.r2, "", saveInFile = NULL, vertices = NULL)
```

# Check Stats of "Positive Control" Pairs.
```{r}
# metabolites_of_interest <- make.names(c(rep("2'-deoxyuridine 5'-triphosphate", 7), rep("citric acid", 11),
#                              rep("malic acid", 6), rep("succinate", 9)))
# genes_of_interest <- make.names(c(c("FURIN", "UMPS", "UPP1", "UPP2", "UCK1", "RRM2", "TYMS"),
#                        c("CNDP1", "APRT", "SLCO2B1", "CLYBL", "BHMT", "PDE5A", "IL4I1", "HS3ST3A1", "HS3ST3B1", "OR10H3"),
#                        c("MDH2", "FH", "ME2", "ME1", "ME3", "MDH1"),
#                        c("TMLHE", "ALKBH2", "PHYH", "P4HA2", "P4HA1", "BBOX1", "PLOD1", "PLOD2", "PLOD3", "P3H1")))
# pairs_of_interest <- paste(genes_of_interest, metabolites_of_interest, sep = "__")
# myres.all <- IntLIM::ProcessResults(inputResults = myres, inputData = inputDataTrain, 
#                                        pvalcutoff = 1, rsquaredCutoff = 0, interactionCoeffPercentile = 0)
# myres.pairs <- myres.all[pairs_of_interest,]
# plot(density(myres.pairs$interaction_coeff))
# lines(density(myres.all$interaction_coeff), col = "red")
# 
# plot(density(myres.pairs$Pval))
# lines(density(myres.all$Pval, col = "red"), col = "red")
# 
# plot(density(myres.pairs$rsquared))
# lines(density(myres.all$rsquared, col = "red"), col = "red")
```

# Perform One-Hot Encoding on Categorical Variables.
```{r}
covar = "cancertype"
encoding <- MultiOmicsGraphPrediction::OneHotEncoding(covar = "cancertype", inputData = inputDataTrain)
covar = encoding$covar
inputDataTrain@sampleMetaData <- encoding$sampleMetaData
```

# Run Pairwise Prediction.
```{r}
# Run pairwise prediction.
pred <- MultiOmicsGraphPrediction::RunPairwisePrediction(inputResults = myres.r2, 
                                                         inputData = inputDataTrain,
                                                         stype = "drug5FU",
                                                         covar = covar,
                                                         independentVarType = 1,
                                                         outcomeType = 2)
#write.csv(pred,file=paste(result_dir, paste0("nci60_5FUPairPredictions.csv"), sep = "\\"))
#hist(pred, breaks = 100)
```

## Project Predictions
```{r plot graph}
# Project graph.
projectedGraph <- MultiOmicsGraphPrediction::ProjectPredictionsOntoGraph(coRegulationGraph = coreg.r2,
                                                                         predictions = pred)
# MultiOmicsGraphPrediction::PlotGraphPredictions(graph = projectedGraph, 
#                                                 inputData = inputDataTrain,
#                                                 stype = "drug5FU")
```

## Compute Importance Metrics.
Future considerations: outcome coefficient, intercept term (we want this to be low).

We assume RaMP has already been installed and that a MySQL version of the RaMP package exists on the user's machine.
__colIdInd__ and __colIdOut__ denote the columns of the analyte metadata that contain the source ID's to input to RaMP.
```{r importance}
#MultiOmicsGraphPrediction::PlotSubspaceClusteringDendrogram(inputData = inputDataTrain, eigStep = 1)
#MultiOmicsGraphPrediction::PlotSubspaceClusteringHeatmap(inputData = inputDataTrain, eigStep = 1)
mysql_password <- unlist(unname(read.csv("mysql_pw.txt", header = FALSE)))
pkg.globals <- RaMP::setConnectionToRaMP(dbname = "ramp", username = "root", conpass = mysql_password,
                                     host = "localhost")
importance <- MultiOmicsGraphPrediction::GetAllImportanceMetrics(predictions = pred, 
                                                                 stype = "drug5FU",
                                                                 inputData = inputDataTrain, 
                                                                 metricList = c("pdf", "interactionpval", "interactioncoef", "pathway", "reaction", "analytecoef", "localerr"),
                                                                 k = 2, eigStep = 1,
                                                                 modelStats = myres.r2,
                                                                 colIdInd = "databaseId",
                                                                 colIdOut = "databaseId")
```

## Create Model Input
```{r model input}
# Create input.
modelInput <- MultiOmicsGraphPrediction::FormatInput(predictionGraphs = projectedGraph, 
                                                     coregulationGraph = coreg.r2,
                                                     inputData = inputDataTrain, 
                                                     stype.class = "numeric",
                                                     stype = "drug5FU",
                                                     edgeTypeList = c("shared.outcome.analyte", "shared.independent.analyte"),
                                                     importance = importance,
                                                     modelProperties = myres.r2,
                                                     covariates = covar)
#saveRDS(modelInput,file=paste(result_dir, paste0("nci60_5FUmodelInput.RDS"), sep = "\\"))
#MultiOmicsGraphPrediction::PlotWeightHeatmap(modelResults = modelResults)
#MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, sampSubset = c("UACC.62", "NCI.ADR.RES", "HL.60.TB.", "UACC.257"), weights = FALSE, stype = "drug5FU", includeLabels = FALSE)
#MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, sampSubset = c("UACC.62", "NCI.ADR.RES", "HL.60.TB.", "UACC.257"), weights = TRUE, stype = "drug5FU", includeLabels = FALSE)
```

# Look at a predictive model.
## Odds Ratio
(2) The final prediction needs to also be a combination of the final subgraphs. We shouldn't be averaging them.
```{r}
modelResults <- MultiOmicsGraphPrediction::InitializeGraphLearningModel(modelInput = modelInput, learningRate = 0.2)
pairsPredAll <- MultiOmicsGraphPrediction::ObtainSubgraphNeighborhoods(modelInput = modelResults@model.input, percentOverlapCutoff = 50)
propagatedGraph <- MultiOmicsGraphPrediction::DoSignificancePropagation(pairs = pairsPredAll, modelResults = modelResults,
                                               verbose = TRUE, makePlots = FALSE, pruningMethod = "odds.ratio")
print(propagatedGraph)
print(MultiOmicsGraphPrediction::CompositePrediction(modelResults = modelResults, pairs = propagatedGraph[[1]]))
print(MultiOmicsGraphPrediction::CompositePrediction(modelResults = modelResults, pairs = propagatedGraph[[2]]))
print(MultiOmicsGraphPrediction::CompositePrediction(modelResults = modelResults, pairs = propagatedGraph[[3]]))
```

## T-Test
(2) The final prediction needs to also be a combination of the final subgraphs. We shouldn't be averaging them.
```{r}
pairsPredAll <- MultiOmicsGraphPrediction::ObtainSubgraphNeighborhoods(modelInput = modelResults@model.input, percentOverlapCutoff = 50)
propagatedGraph <- MultiOmicsGraphPrediction::DoSignificancePropagation(pairs = pairsPredAll, modelResults = modelResults,
                                               verbose = TRUE, makePlots = FALSE, pruningMethod = "error.t.test")
print(propagatedGraph)
print(MultiOmicsGraphPrediction::CompositePrediction(modelResults = modelResults, pairs = propagatedGraph[[1]]))
```

# Learn Weights.
```{r}
modelInput <- MultiOmicsGraphPrediction::FormatInput(predictionGraphs = projectedGraph, 
                                                     coregulationGraph = coreg.r2,
                                                     inputData = inputDataTrain, 
                                                     stype.class = "numeric",
                                                     stype = "drug5FU",
                                                     edgeTypeList = c("shared.outcome.analyte", "shared.independent.analyte"),
                                                     importance = importance,
                                                     modelProperties = myres.r2,
                                                     covariates = covar,
                                                     predictorBounds = c(-5,5))
modelResults <- MultiOmicsGraphPrediction::InitializeGraphLearningModel(modelInput = modelInput, learningRate = 0.2)
optimalModel <- MultiOmicsGraphPrediction::OptimizeImportanceCombo(modelResults = modelResults, verbose = TRUE)
```

## Calculate information gain for when all terms are included in a composite predictor.
Run the first iteration of composite prediction.
```{r}
# pairsPredAll <- MultiOmicsGraphPrediction::ObtainSubgraphNeighborhoods(modelInput = modelResults@model.input, percentOverlapCutoff = 50)
# propagatedGraph <- MultiOmicsGraphPrediction::DoSignificancePropagation(pairs = pairsPredAll, modelResults = modelResults,
#                                                 verbose = TRUE, makePlots = TRUE)
```
# Line Graphs (Before and After)
```{r}
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, sampSubset = c("UACC.62", "NCI.ADR.RES", "HL.60.TB.", "UACC.257"), weights = FALSE, stype = "drug5FU", includeLabels = FALSE)
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, subset = unlist(propagatedGraph), sampSubset = c("UACC.62", "NCI.ADR.RES", "HL.60.TB.", "UACC.257"), weights = FALSE, stype = "drug5FU", includeLabels = FALSE)
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, sampSubset = c("UACC.62", "NCI.ADR.RES", "HL.60.TB.", "UACC.257"), weights = TRUE, stype = "drug5FU", includeLabels = FALSE)
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, subset = unlist(propagatedGraph), sampSubset = c("UACC.62", "NCI.ADR.RES", "HL.60.TB.", "UACC.257"), weights = TRUE, stype = "drug5FU", includeLabels = FALSE)
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, sampSubset = "NCI.ADR.RES", weights = TRUE, stype = "drug5FU")
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, subset = unlist(propagatedGraph), sampSubset = "NCI.ADR.RES", weights = TRUE, stype = "drug5FU")
# 
# png(paste(getwd(), "NCI.ADR.RES.full.png", sep = "/"), res = 300, width = 5, height = 4, units = "in")
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, sampSubset = "NCI.ADR.RES", weights = FALSE, stype = "drug5FU", includeLabels = FALSE)
# dev.off()
# png(paste(getwd(), "NCI.ADR.RES.pruned.png", sep = "/"), res = 300, width = 5, height = 4, units = "in")
# MultiOmicsGraphPrediction::PlotLineGraph(modelResults = modelResults, subset = unlist(propagatedGraph), sampSubset = "NCI.ADR.RES", weights = TRUE, stype = "drug5FU", includeLabels = FALSE)
# dev.off()
```

# Plot Results
```{r}
# predPairs1 <- MultiOmicsGraphPrediction::CompositePrediction(pairs = prunedAgain[[1]],
#                                                                          IntLIMresults = myres,
#                                                                          modelResults = modelResults,
#                                                                          covar = covar,
#                                                                          inputData = inputDataTrain)
# predPairs2 <- MultiOmicsGraphPrediction::CompositePrediction(pairs = prunedAgain[[2]],
#                                                                          IntLIMresults = myres,
#                                                                          modelResults = modelResults,
#                                                                          covar = covar,
#                                                                          inputData = inputDataTrain)
# plot(inputDataTrain@sampleMetaData$drug5FU, predPairs1, xlab = "True Value", ylab = "Predicted Value (Composite Model 1)", ylim = range(inputDataTrain@sampleMetaData$drug5FU))
# plot(inputDataTrain@sampleMetaData$drug5FU, predPairs2, xlab = "True Value", ylab = "Predicted Value (Composite Model 2)", ylim = range(inputDataTrain@sampleMetaData$drug5FU))
# plot(inputDataTrain@sampleMetaData$drug5FU, (predPairs2 + predPairs1) / 2, xlab = "True Value", ylab = "Predicted Value (Composite Model 2)", ylim = range(inputDataTrain@sampleMetaData$drug5FU))
# plot(inputDataTrain@sampleMetaData$drug5FU, predPairs1, xlab = "True Value", ylab = "Predicted Value (Composite Model 1)", ylim = c(0,10))
# plot(inputDataTrain@sampleMetaData$drug5FU, predPairs2, xlab = "True Value", ylab = "Predicted Value (Composite Model 2)", ylim = c(0,10))
# plot(inputDataTrain@sampleMetaData$drug5FU, (predPairs2 + predPairs1) / 2, xlab = "True Value", ylab = "Predicted Value (Composite Model 2)", ylim = c(0,10))
```