
---
title: "NCI-60 5-FU Demo"
author: "Tara Eicher"
date: "4/01/2022"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{MultiOmicsGraphPrediction NCI60 5FU}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# Set up.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
result_dir <- "~\\Ensemble_5fu_vignette_results"
dir.create(result_dir)
```

# Install the packages.
```{r eval = FALSE}
if(!require("devtools")){
  install.packages("devtools")
}
library("devtools")
if(!require("MultiOmicsGraphPrediction")){
  install_github("ncats/MultiOmicsGraphPrediction")
}
library(MultiOmicsGraphPrediction)
if(!require("IntLIM")){
  install_github("ncats/IntLIM")
}
library(IntLIM)
```

# Import data files.
Read in the full data set.
```{r}
#dir <- system.file("extdata", package="MultiOmicsGraphPrediction", mustWork=TRUE)
dir <- "/Volumes/eichertd$/Ensemble_prediction/inst/extdata"
# Import training data.
csvfile <- file.path(dir, "nci60_5FU.full.csv")
inputData <- IntLIM::ReadData(csvfile, class.feat = list(drugscore = "numeric",
                                                                   cancertype = "factor",
                                                                   drug5FU = "numeric"))
```

# Run IntLIM
We observe that R^2 values range between 0.3 and 0.8 and that p-values are generally
high. In this case, we will use only the R^2 value to filter the data. This is
because pairs may be important to prediction even if their association is not
phenotype-dependent.
```{r}
myres <- IntLIM::RunIntLim(inputData = inputData,stype="drug5FU",
                                 save.covar.pvals = TRUE, 
                                 outcome = 2, suppressWarnings = TRUE,
                                 independent.var.type = 1, 
                                 continuous = TRUE, covar = "cancertype")
IntLIM::DistPvalues(myres, adjusted = FALSE)
IntLIM::DistRSquared(myres)
IntLIM::InteractionCoefficientGraph(inputResults = myres,
                                    interactionCoeffPercentile = 0.9,
                                    percentageToPlot = 1)
IntLIM::pvalCoefVolcano(inputResults = myres, inputData = inputData,
                        pvalcutoff = 0.99)
plot(density(myres@interaction.pvalues), xlab = "Interaction Term Unadjusted p-value", main = "")
plot(density(myres@interaction.adj.pvalues), xlab = "Interaction Term Adjusted p-value", main = "")
plot(density(myres@interaction.coefficients), xlab = "Interaction Term Coefficient", main = "")
plot(density(myres@model.rsquared), xlab = "Model R-squared", main = "")
saveRDS(myres, "/Volumes/eichertd$/Ensemble_prediction/vignettes/myres_NCI60_full.RDS")
```

# Filter Results and Create Co-Regulation Graph
After filtering by the R^2 value at the peak of the distribution (0.38), we obtain
2,076,567 pairs that pass the threshold. This is out of an original 5,018,373 possible pairs.

However, we note that only cysteine and succinate are involved in pairs that pass the
R^2 threshold. We would also like to consider the role of dUTP, citric acid, and malic acid.
For this reason, we proceed using all pairs rather than only those filtered using R^2.
```{r}
myres.all <- IntLIM::ProcessResults(inputResults = myres, inputData = inputData, 
                                       pvalcutoff = 1, rsquaredCutoff = 0, interactionCoeffPercentile = 0)
myres.r2.density <- density(myres.all$rsquared)
peak.r2 <- myres.r2.density$x[which.max(myres.r2.density$y)]
print(peak.r2)
myres.r2 <- IntLIM::ProcessResults(inputResults = myres, inputData = inputData, 
                                       pvalcutoff = 1, rsquaredCutoff = peak.r2)
# Build graph.
coreg.r2 <- MultiOmicsGraphPrediction::BuildCoRegulationGraph(myres.r2)

# Check whether metabolites and genes of interest are present.
truePositiveMetabolites <- unlist(read.table("/Volumes/eichertd$/Ensemble_prediction/inst/extdata/truePositiveMetabolites.txt",
            sep = "\n", header = FALSE)$V1)
truePositiveGenes <- unlist(read.table("/Volumes/eichertd$/Ensemble_prediction/inst/extdata/truePositiveGenes.txt",
            sep = "\n", header = FALSE)$V1)
trueNegativeGenes <- unlist(read.table("/Volumes/eichertd$/Ensemble_prediction/inst/extdata/trueNegativeGenes.txt",
            sep = "\n", header = FALSE)$V1)
print(paste(length(intersect(as.character(myres.r2$Analyte2), truePositiveMetabolites)), length(truePositiveMetabolites)))
print(paste(length(intersect(as.character(myres.r2$Analyte1), truePositiveGenes)), length(truePositiveGenes)))
print(paste(length(intersect(as.character(myres.r2$Analyte1), trueNegativeGenes)), length(trueNegativeGenes)))
saveRDS(myres.r2, "/Volumes/eichertd$/Ensemble_prediction/vignettes/myres_NCI60_full_r2.RDS")
saveRDS(coreg.r2, "/Volumes/eichertd$/Ensemble_prediction/vignettes/coreg_NCI60_full_r2.RDS")
```

# Perform One-Hot Encoding on Categorical Variables.

We perform one-hot encoding on the categorical variables.
```{r}
covar = "cancertype"
encoding <- MultiOmicsGraphPrediction::OneHotEncoding(covar = "cancertype", inputData = inputData)
covar = encoding$covar
inputData@sampleMetaData <- encoding$sampleMetaData
saveRDS(inputData, "/Volumes/eichertd$/Ensemble_prediction/vignettes/inputData_NCI60_full_r2.RDS")
```

# Run Pairwise Prediction.

Next, we predict the outcomes. As we can observe from the histogram plots, predictions
from each pair are not necessarily centered around their true values. We therefore
anticipate that the PDF feature will be relatively unimportant to the final predictions.
```{r}
myres.r2 <- readRDS("/Volumes/eichertd$/Ensemble_prediction/vignettes/myres_NCI60_full_r2.RDS")
# Run pairwise prediction.
pred <- MultiOmicsGraphPrediction::RunPairwisePrediction(inputResults = myres.r2, 
                                                         inputData = inputData,
                                                         stype = "drug5FU",
                                                         covar = covar,
                                                         independentVarType = 1,
                                                         outcomeType = 2)
saveRDS(pred,file="/Volumes/eichertd$/Ensemble_prediction/vignettes/nci60_5FUPairPredictions.full.RDS")
for(samp in rownames(pred)){
  hist(pred[samp,], main = paste(samp, inputData@sampleMetaData[samp,"drug5FU"]), xlim = c(0,10), breaks = 1000000)
}
```

## Project Predictions
```{r plot graph}
#pred <- readRDS(file="/Volumes/eichertd$/Ensemble_prediction/vignettes/nci60_5FUPairPredictions.full.RDS")
coreg.r2 <- readRDS("/Volumes/eichertd$/Ensemble_prediction/vignettes/coreg_NCI60_full_r2.RDS")
# Project graph.
projectedGraph <- MultiOmicsGraphPrediction::ProjectPredictionsOntoGraph(coRegulationGraph = coreg.r2,
                                                                         predictions = pred)
saveRDS(projectedGraph, "/Volumes/eichertd$/Ensemble_prediction/vignettes/projectedGraph_NCI60_full_r2.RDS")
```

## Compute Metafeatures.
We assume RaMP has already been installed and that a MySQL version of the RaMP package exists on the user's machine.
__colIdInd__ and __colIdOut__ denote the columns of the analyte metadata that contain the source ID's to input to RaMP.

For most of the metafeatures, the distribution of values is fairly uniform. The exception to this is the interaction p-value
metafeature, which peaks at a p-value of 0.4 (not particularly significant). None of the pairs evaluated here share a
reaction in RaMP, although many do share a pathway.

Because clusters are not apparent within the subspace clustering heatmap, it is expected that the local error of each
predictor will mirror the global error.
```{r importance}
MultiOmicsGraphPrediction::PlotSubspaceClusteringDendrogram(inputData = inputData, eigStep = 1)
MultiOmicsGraphPrediction::PlotSubspaceClusteringHeatmap(inputData = inputData, eigStep = 1)
mysql_password <- unlist(unname(read.csv("mysql_pw.txt", header = FALSE)))
pkg.globals <- RaMP::setConnectionToRaMP(dbname = "ramp", username = "root", conpass = mysql_password,
                                     host = "localhost")
metaFeatures <- MultiOmicsGraphPrediction::GetMetaFeatures(predictions = pred,
                                                                 stype = "drug5FU",
                                                                 inputData = inputData,
                                                                 metaFeatureList = c("pdf", "interactionpval", "interactioncoef", "pathway", "reaction", "analytecoef", "localerr"),
                                                                 k = 2, eigStep = 1,
                                                                 modelStats = myres.r2,
                                                                 colIdInd = "databaseId",
                                                                 colIdOut = "databaseId")
# metaFeatures <- MultiOmicsGraphPrediction::GetMetaFeatures(predictions = pred,
#                                                                  stype = "drug5FU",
#                                                                  inputData = inputData,
#                                                                  metaFeatureList = c("pdf", "pathway", "reaction", "localerr"),
#                                                                  k = 2, eigStep = 1,
#                                                                  modelStats = myres.r2,
#                                                                  colIdInd = "databaseId",
#                                                                  colIdOut = "databaseId")
# metaFeaturesX <- MultiOmicsGraphPrediction::GetMetaFeatures(predictions = pred,
#                                                                  stype = "drug5FU",
#                                                                  inputData = inputData,
#                                                                  metaFeatureList = c("interactionpval", "interactioncoef", "analytecoef"),
#                                                                  k = 2, eigStep = 1,
#                                                                  modelStats = myres.r2,
#                                                                  colIdInd = "databaseId",
#                                                                  colIdOut = "databaseId")
# metaFeatures$interactionpval <- metaFeaturesX$interactionpval
# metaFeatures$interactioncoef <- metaFeaturesX$interactioncoef
# metaFeatures$analytecoef <- metaFeaturesX$analytecoef
hist(metaFeatures$pdf)
hist(metaFeatures$localerr)
hist(metaFeatures$pathway)
hist(metaFeatures$reaction)
hist(metaFeatures$interactionpval)
hist(metaFeatures$interactioncoef)
hist(metaFeatures$analytecoef)
saveRDS(metaFeatures, "/Volumes/eichertd$/Ensemble_prediction/vignettes/metaFeatures_NCI60_full_r2.RDS")
```

## Run a single Significance Propagation with the initial weights.
The pairs that we expect to find in the significance model include the following:
- For dUTP: FURIN, UMPS, UPP1, UPP2, UCK1, RRM2, TYMS
- For citric acid: CNDP1, APRT, SLCO2B1, CLYBL, BHMT, PDE5A, IL4I1, HS3ST3A1, HS3ST3B1, and OR10H3
- For malic acid: MDH2, FH, ME2, ME1, ME3, and MDH1
- For succinate: TMLHE, ALKBH2, PHYH, P4HA2, P4HA1, BBOX1, PLOD1, PLOD2, PLOD3, and P3H1

The pairs learned using only a single propagation are as follows:
- For malic acid: FH, CLYBL, HS3ST3A1, MDH1, PLOD2, UPP1, and CNDP1
- For cysteine: CNDP1
- For succinate: CNDP1
- For citric acid: CNDP1

Only two of these pairs overlap with the expected pairs. Furthermore, the t-scores
are negative. This indicates that the overall error is greater than if the mean
value is used for prediction. Finally, the predictions are far from the actual
true values.
```{r model input}
metaFeatures <- readRDS("/Volumes/eichertd$/Ensemble_prediction/vignettes/metaFeatures_NCI60_full_r2.RDS")
projectedGraph <- readRDS("/Volumes/eichertd$/Ensemble_prediction/vignettes/projectedGraph_NCI60_full_r2.RDS")

modelInput <- MultiOmicsGraphPrediction::FormatInput(predictionGraphs = projectedGraph, 
                                                     coregulationGraph = coreg.r2,
                                                     inputData = inputData, 
                                                     stype.class = "numeric",
                                                     stype = "drug5FU",
                                                     edgeTypeList = c("shared.outcome.analyte", "shared.independent.analyte"),
                                                     metaFeatures = metaFeatures,
                                                     modelProperties = myres.r2,
                                                     outcome = 2,
                                                     independent.var.type = 1)
modelResults <- MultiOmicsGraphPrediction::InitializeGraphLearningModel(modelInput = modelInput, learningRate = 0.2)
pairsPredAll <- MultiOmicsGraphPrediction::ObtainSubgraphNeighborhoods(modelInput = modelResults@model.input)
saveRDS(pairsPredAll, "/Volumes/eichertd$/Ensemble_prediction/vignettes/pairsPredAll_NCI60_full_r2.RDS")
propagatedGraph <- MultiOmicsGraphPrediction::DoSignificancePropagation(pairs = pairsPredAll, modelResults = modelResults,
                                               verbose = TRUE, modelRetention = "stringent")
print(propagatedGraph)
print(MultiOmicsGraphPrediction::CompositePrediction(modelResults = modelResults, pairs = propagatedGraph[[1]]))
```

## Run a single Significance Propagation with prediction cutoffs.
The pairs that we expect to find in the significance model include the following:
- For dUTP: FURIN, UMPS, UPP1, UPP2, UCK1, RRM2, TYMS
- For citric acid: CNDP1, APRT, SLCO2B1, CLYBL, BHMT, PDE5A, IL4I1, HS3ST3A1, HS3ST3B1, and OR10H3
- For malic acid: MDH2, FH, ME2, ME1, ME3, and MDH1
- For succinate: TMLHE, ALKBH2, PHYH, P4HA2, P4HA1, BBOX1, PLOD1, PLOD2, PLOD3, and P3H1

The pairs learned using only a single propagation include the dUTP pairs.

Two of these pairs overlap with the expected pairs (dUTP, FURIN), and (dUTP, RRM2).
Notably, none of the other dUTP pairs passed the R^2 threshold, so these were not
considered. 
```{r model input}
modelInput <- MultiOmicsGraphPrediction::FormatInput(predictionGraphs = projectedGraph, 
                                                     coregulationGraph = coreg.r2,
                                                     inputData = inputDataTrain, 
                                                     stype.class = "numeric",
                                                     stype = "drug5FU",
                                                     edgeTypeList = c("shared.outcome.analyte", "shared.independent.analyte"),
                                                     metaFeatures = metaFeatures,
                                                     modelProperties = myres.r2,
                                                     outcome = 2,
                                                     independent.var.type = 1)
modelResults <- MultiOmicsGraphPrediction::InitializeGraphLearningModel(modelInput = modelInput, learningRate = 0.2, iterations = 100)
pairsPredAll <- MultiOmicsGraphPrediction::ObtainSubgraphNeighborhoods(modelInput = modelResults@model.input, percentOverlapCutoff = 50)

maxCutoff <- 1.1 * max(modelInput@input.data@sampleMetaData[,"drug5FU"])
minCutoff <- 0.9 * min(modelInput@input.data@sampleMetaData[,"drug5FU"])
propagatedGraph <- MultiOmicsGraphPrediction::DoSignificancePropagation(pairs = pairsPredAll, modelResults = modelResults,
                                               verbose = TRUE, modelRetention = "stringent", minCutoff = minCutoff, maxCutoff = maxCutoff, useCutoff = TRUE)
predictions <- MultiOmicsGraphPrediction::CompositePrediction(modelResults = modelResults, pairs = propagatedGraph[[1]], minCutoff = minCutoff,
                                                     maxCutoff = maxCutoff, useCutoff = TRUE)
true <- modelInput@input.data@sampleMetaData[,"drug5FU"]
print(propagatedGraph)
plot(predictions, true, xlab = "Predicted", ylab = "True")
print(ComputeRMSE(predictions, true))
```

# Learn Metafeature Weights.

In this case, errors are very high during model tuning, making convergence impractical.
This is likely because many of the predictions fall outside of the true range of the data.
Capping the predictions given the training data may help to mitigate this issue.
```{r}
# Initialize parameters.
metaFeatureList <- c("pdf", "interactionpval", "interactioncoef", "pathway", "reaction", "analytecoef", "localerr")
stype <- "drug5FU"
outcomeType <- 2
independentVarType <- 1

# Run prediction.
modelResults <- MultiOmicsGraphPrediction::DoModelSetup(inputData = inputDataTrain,
                                stype = stype,
                                rsquaredCutoff = 0,
                                outcomeType = outcomeType,
                                independentVarType = independentVarType,
                                continuous = TRUE,
                                metaFeatureList = metaFeatureList,
                                learningRate = 0.01,
                                maxIterations = 100)
optimalModel <- MultiOmicsGraphPrediction::OptimizeMetaFeatureCombo(modelResults = modelResults, verbose = TRUE)

# Predict testing values.
predictions <- DoTestSetupAndPrediction(inputDataTest = inputDataTest, 
                                        model = optimalModel)
print(optimalModel@pairs)
print(predictions)
print(inputDataTest@sampleMetaData$drug5FU)
print(MultiOmicsGraphPrediction::ComputeRMSE(true.Y = inputDataTest@sampleMetaData$drug5FU,
                                             pred.Y = predictions))
plot(predictions, inputDataTest@sampleMetaData$drug5FU, xlab = "Predicted 5-FU Drug Score", ylab = "True 5-FU Drug Score")
```

# Learn Metafeature Weights with Prediction Cutoffs.
```{r}
# Initialize parameters.
metaFeatureList <- c("pdf", "interactionpval", "interactioncoef", "pathway", "reaction", "analytecoef", "localerr")
stype <- "drug5FU"
outcomeType <- 2
independentVarType <- 1

# Run prediction.
modelResults <- DoModelSetup(inputData = inputDataTrain,
                                stype = stype,
                                outcomeType = outcomeType,
                                independentVarType = independentVarType,
                                continuous = TRUE,
                                metaFeatureList = metaFeatureList,
                             rsquaredCutoff = 0,
                             maxIterations = 100,
                                learningRate = 0.01)
optimalModel <- MultiOmicsGraphPrediction::OptimizeMetaFeatureCombo(modelResults = modelResults, useCutoff = TRUE)

# Predict testing values.
predictions <- DoTestSetupAndPrediction(inputDataTest = inputDataTest, 
                                        model = optimalModel, 
                                        useCutoff = TRUE)
print(optimalModel@pairs)
print(predictions)
print(inputDataTest@sampleMetaData$drug5FU)
print(MultiOmicsGraphPrediction::ComputeRMSE(true.Y = inputDataTest@sampleMetaData$drug5FU,
                                             pred.Y = predictions))
plot(predictions, inputDataTest@sampleMetaData$drug5FU, xlab = "Predicted 5-FU Drug Score", ylab = "True 5-FU Drug Score")
```

# Random Forest
```{r}
# Install packages.
if(!require("caret")){
  install.packages("caret")
}
if(!require("randomForest")){
  install.packages("randomForest", dependencies = TRUE)
}
library(randomForest)
set.seed(1234)
```

```{r}
# Set up Data.
train.X.all <- as.matrix(cbind(t(inputDataTrain@analyteType1),t(inputDataTrain@analyteType2)))
test.X.all <- as.matrix(cbind(t(inputDataTest@analyteType1),t(inputDataTest@analyteType2)))
train.Y.all <- inputDataTrain@sampleMetaData[, "drug5FU"]
test.Y.all <- inputDataTest@sampleMetaData[, "drug5FU"]

# Run Random Forest
control <- caret::trainControl(method='none')
mtry <- sqrt(ncol(train.X.all))
tunegrid <- expand.grid(.mtry=mtry)
rf_fit <- caret::train(x = train.X.all, y = c(train.Y.all),
                method='rf',
                metric='RMSE',
                tuneGrid=tunegrid,
                trControl=control)
rf_models <- rf_fit[which(names(rf_fit) != "trainingData")]
rf_importance <- caret::varImp(rf_fit, scale = FALSE)
rf_test_predict <- predict(rf_fit, test.X.all)
rf_test_true <- test.Y.all
rf_rmse <- ComputeRMSE(test.Y.all, rf_test_predict)
print(rownames(rf_importance$importance)[order(-1 * rf_importance$importance$Overall)])
print(paste(sample, rf_test_predict, rf_test_true))
print(rf_rmse)
plot(rf_test_predict, rf_test_true, xlab = "Predicted 5-FU Drug Score", ylab = "True 5-FU Drug Score")
```


# LASSO
```{r}
# Install packages.
if(!require("glmnet")){
  install.packages("glmnet", dependencies = TRUE)
}
```

```{r}
# Set up Data.
train.X.all <- as.matrix(cbind(t(inputDataTrain@analyteType1),t(inputDataTrain@analyteType2)))
test.X.all <- as.matrix(cbind(t(inputDataTest@analyteType1),t(inputDataTest@analyteType2)))
train.Y.all <- inputDataTrain@sampleMetaData[, "drug5FU"]
test.Y.all <- inputDataTest@sampleMetaData[, "drug5FU"]

# Run LASSO
control <- caret::trainControl(method='none')
tunegrid <- expand.grid(alpha = 1, lambda = 0)
lasso_fit <- caret::train(x = train.X.all, y = c(train.Y.all),
                   method='glmnet',
                   metric='RMSE',
                   tuneGrid=tunegrid,
                   trControl=control)
lasso_models <- lasso_fit[which(names(lasso_fit) != "trainingData")]
lasso_importance <- coef(lasso_fit$finalModel)
lasso_test_predict <- predict(lasso_fit, test.X.all)
lasso_test_true <- test.Y.all
lasso_rmse <- ComputeRMSE(test.Y.all, lasso_test_predict)
gplots::heatmap.2(as.matrix(lasso_importance), trace = "none", scale = "none")
print(paste(sample, lasso_test_predict, lasso_test_true))
print(lasso_rmse)
plot(lasso_test_predict, lasso_test_true, xlab = "Predicted 5-FU Drug Score", ylab = "True 5-FU Drug Score")
```
